Efficient ancestry and mutation simulation with msprime 1.0

### 2021-11-19 paper.tex
Cut down Demography section

\subsection*{Demography}
\label{sec-demography}
% TODO should we put in some citations here for the models? Hardly seems
% necessary.
One of the key applications of population genetic simulations is to generate
data for complex demographies. Beyond idealised cases such as stepping-stone or
island models, or specialised cases such as isolation-with-migration models,
analytical results are rarely possible. Simulation is therefore integral to the
development and evaluation of methods for demographic inference. The demography
model in \msprime\ is directly derived from the approach used in \ms, and
supports an arbitrary number of randomly mating populations exchanging migrants
at specified rates. A range of demographic events are supported, which allow for
varying population sizes and growth rates, changing migration rates over time,
as well as population splits, admixtures and pulse migrations.
The location of sampled lineages
can be tracked through time in as much detail as required: each tree node is
automatically associated with the population in which it arose, the location of
lineages can be recorded at any given time via census events, or every lineage
migration can be recorded.
Large demographic models can be simulated efficiently in version \msprime\ 1.0,
since we only consider populations that contain lineages and
have non-zero migration rates when generating migration event
waiting times. This is a considerable improvement
over version 0.x, which scaled quadratically with the number of populations.

A major change for \msprime\ 1.0 is the introduction of the new Demography API,
designed to address a design flaw in the \msprime\ 0.x interface which led to
a number of avoidable errors in downstream
simulations~\citep{ragsdale2020lessons}. Briefly, the 0.x API required three
separate parameters be provided to the \texttt{simulate} function to describe a
demographic model, making it easy to accidentally omit information. The 1.0 API
resolves this issue by creating a new \texttt{Demography} class, which
encapsulates all information about the demographic model, and fully decouples
the definition from other simulation details. An instance of this class is then
provided as a parameter to the new \texttt{sim\_ancestry} function,
substantially reducing the potential for error.
Another improvement over the
0.x APIs is the introduction of explicit population split and admixture events,
and a population state machine that ensures that lineages cannot migrate into
(or be sampled from) inactive populations. This demography model is compatible
with the Demes standard~\citep{gower2022demes}, and the \texttt{Demography}
class supports importing and exporting Demes models.
Models previously constructed using the 0.x API can be seamlessly imported into
the \texttt{Demography} class, and we also support importing demographic
models from Newick species trees and the output of programs
like *BEAST~\citep{heled2009bayesian}.

The \texttt{DemographyDebugger} provides detailed information about
demographic models as well as numerical methods to make predictions
about these models. For example, we can compute
the coalescence rates for two or more lineages
drawn from populations at specified times in the past,
which can be inverted to obtain the
``inverse instantaneous coalescence rate'' of \citet{chikhi2018iicr}.
Many popular approaches in population genetics use the distribution of
coalescence rates between pairs of lineages in one or more populations to infer
effective population sizes over
time~\citep{li2011inference,sheehan2013estimating,schiffels2014inferring} or
split times and subsequent migration rates between
populations~\citep{wang2020tracking}.
These numerical methods provide a valuable
ground-truth when evaluating such inference methods, as illustrated by
\cite{adrion2020community}.

>>>>

\subsection*{Demography}
\label{sec-demography}
One of the key applications of population genetic simulations is to generate
data for complex demographies. Beyond idealised cases such as stepping-stone or
island models, or specialised cases such as isolation-with-migration models,
analytical results are rarely possible. Simulation is therefore integral to the
development and evaluation of methods for demographic inference. The demography
model in \msprime\ is directly derived from the approach used in \ms, and
supports an arbitrary number of randomly mating populations exchanging migrants
at specified rates. A range of demographic events are supported, which allow for
varying population sizes and growth rates, changing migration rates over time,
as well as population splits, admixtures and pulse migrations.

A major change for \msprime\ 1.0 is the introduction of the new Demography API,
designed to address a design flaw in the \msprime\ 0.x interface which led to
avoidable errors in downstream simulations~\citep{ragsdale2020lessons}.
The new API is more user-friendly, providing the ability, for example,
to refer to populations by name rather than their integer identifiers.
We also provide numerical methods to compute
the coalescence rates for two or more lineages
which can be inverted to obtain the
``inverse instantaneous coalescence rate'' of \citet{chikhi2018iicr}.
Many popular approaches in population genetics use the distribution of
coalescence rates between pairs of lineages to infer
effective population sizes over
time~\citep{li2011inference,sheehan2013estimating,schiffels2014inferring} or
split times and subsequent migration rates between
populations~\citep{wang2020tracking}.
These numerical methods provide a valuable
ground-truth when evaluating such inference methods, as illustrated by
\cite{adrion2020community}.
